---
title: "p8105_hw1_tl3279"
author: "Tianqi Li"
date: "2024-09-16"
output: github_document
---

Load necessary packages, and setup a temporary directory to avoid R Markdown to automatically save the plot for question 1.
```{r echo = FALSE, message = FALSE}
knitr::opts_chunk$set(fig.path = tempdir())
library(ggplot2)
library(tidyverse)
```

## Problem 1

Load the dataset and describe it.

```{r}
data("penguins", package = "palmerpenguins")
summary(penguins)
nrow(penguins)
ncol(penguins)
mean(penguins$flipper_length_mm, na.rm = TRUE)
```
The variables in this dataset include species (Adelie, Chinstrap, or Gentoo), island (Biscoe, Dream, or Torgersen),bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex (female, male, or NA), and year (2007, 2008, 0r 2009).

The dataset has 344 rows and 8 columns (344 observations of 8 variables).

The mean flipper length equals to 200.9mm.

Plot and save the plot
```{r warning = FALSE}
scatter_plot <- ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
print(scatter_plot)
ggsave("flipper_vs_bill.png", scatter_plot,path = getwd())
```

## Problem 2
Create the data frame
```{r}
my_df <- tibble(
  numeric_var = rnorm(10),
  logical_var = numeric_var > 0,
  character_var = sample(letters, 10),
  factor_var = factor(sample(c("Level1", "Level2", "Level3"), 10, replace = TRUE))
)
print(my_df)
```

Calculate the mean for each variable.
```{r warning = FALSE}
mean_numeric <- mean(pull(my_df, numeric_var))
mean_logical <- mean(pull(my_df, logical_var))
mean_character <- mean(pull(my_df, character_var))
mean_factor <- mean(pull(my_df, factor_var), na.rm = TRUE)
```

And print out the results.

```{r}
print(mean_numeric)
print(mean_logical)
print(mean_character)
print(mean_factor)
```

The mean works for the numerical variable and the logical variable, but not for the character variable or the factor variable.

```{r eval = FALSE}
numeric_logical <- as.numeric(pull(my_df, logical_var))
numeric_character <- as.numeric(pull(my_df, character_var))
numeric_factor <- as.numeric(pull(my_df, factor_var))
print(numeric_logical)
print(numeric_character)
print(numeric_factor)
mean(numerica_factor)
```
as.numeric(logical_var) converts TRUE to 1 and FALSE to 0 which allows to take the mean.
as.numeric(character_var) will return "NA", since characters can't be directly converted to numbers.
as.numeric(factor_var) converts factors to their integer codes (e.g., "Level1" = 1, "Level2" = 2, etc.). But the mean function cannot directly compute a factor variable so "NA" is returned. We can compute the mean bu using mean(as.numerica(factor_var)).

